Probabilistic Catalogs from Crowded Star Fields

### 2012-08-03 paper/ms.tex
Deriving again

Consider the complementary CDF:
\begin{eqnarray}
G(x) = P(X \geq x)
\end{eqnarray}
The complementary CDF for a broken Pareto is:
\begin{eqnarray}
G(x) &=&
\left\{
\begin{array}{lcr}
1, & & x < x_1 \\
\left(\frac{x_1}{x}\right)^{\alpha_1}, & & x_1 \leq x \leq x_2 \\
K\left(\frac{x_2}{x}\right)^{\alpha_2}, & & x > x_2
\end{array}
\right.
\end{eqnarray}
Since we want the CDF to be continuous (to avoid delta functions in the PDF!)
then the value of $K$ must be set by:
\begin{eqnarray}
\left(\frac{x_1}{x_2}\right)^{\alpha_1} &=& K\left(\frac{x_2}{x_2}\right)^{\alpha_2} \\
\implies K &=& \left(\frac{x_1}{x_2}\right)^{\alpha_1}.
\end{eqnarray}
Therefore the complementary CDF is:
\begin{eqnarray}
G(x) &=&
\left\{
\begin{array}{lcr}
1, & & x < x_1 \\
\left(\frac{x_1}{x}\right)^{\alpha_1}, & & x_1 \leq x \leq x_2 \\
\left(\frac{x_1}{x_2}\right)^{\alpha_1}\left(\frac{x_2}{x}\right)^{\alpha_2}, & & x > x_2
\end{array}
\right.
\end{eqnarray}
From this we can obtain the standard CDF:
\begin{eqnarray}
F(x) &=& P(X \leq x) \\
&=& 1 - G(x)\\
&=& \left\{
\begin{array}{lcr}
0, & & x < x_1 \\
1 - \left(\frac{x_1}{x}\right)^{\alpha_1}, & & x_1 \leq x \leq x_2 \\
1 - \left(\frac{x_1}{x_2}\right)^{\alpha_1}\left(\frac{x_2}{x}\right)^{\alpha_2}, & & x > x_2
\end{array}
\right.
\end{eqnarray}
and its inverse (defined on $[0, 1]$):
\begin{eqnarray}
F^{-1}(x) 
&=& \left\{
\begin{array}{lcr}
x_1(1-x)^{-\frac{1}{\alpha_1}}, & & 0 \leq x \leq 1 - \left(\frac{x_1}{x_2}\right)^{\alpha_1} \\
x_2\left((1-x)\left(\frac{x_2}{x_1}\right)^{\alpha_1}\right)^{-\frac{1}{\alpha_2}}, & & 1 - \left(\frac{x_1}{x_2}\right)^{\alpha_1} \leq x \leq 1
\end{array}
\right.
\end{eqnarray}

The probability density function (PDF) can be obtained by differentiation:
\begin{eqnarray}
f(x) &=& \frac{d}{dx} F(x) \\
&=& \left\{
\begin{array}{lcr}
0, & & x < x_1 \\
\left[\alpha_1 x_1^{\alpha_1}\right]x^{-\alpha_1 - 1}, & & x_1 \leq x \leq x_2 \\
\left[\alpha_2 x_1^{\alpha_1} x_2^{\alpha_2 - \alpha_1}\right]x^{-\alpha_2 - 1}, & & x > x_2
\end{array}
\right.
\end{eqnarray}
The normalization constants are the factors inside square brackets.

>>>>

The second condition is the total probability must be 1:

\begin{eqnarray}
\int_{x_1}^{x_2} Z_1^{-1} x^{-\alpha_1 - 1} \, dx
+
\int_{x_2}^\infty Z_2^{-1} x^{-\alpha_2 - 1} \, dx
&=& 1 \\
Z_1^{-1}\alpha_1^{-1}\left[x_1^{-\alpha_1} - x_2^{-\alpha_1}\right]
+
Z_2^{-1}\alpha_2^{-1}x_2^{-\alpha_2}
&=& 1 \\
Z_1^{-1}\alpha_1^{-1}\left[x_1^{-\alpha_1} - x_2^{-\alpha_1}\right]
+
Z_1^{-1}x_2^{\alpha_2-\alpha_1}\alpha_2^{-1}x_2^{-\alpha_2}
&=& 1
\end{eqnarray}

\begin{eqnarray}
\implies
Z_1 &=& \alpha_1^{-1}\left[x_1^{-\alpha_1} - x_2^{-\alpha_1}\right]
+
x_2^{-\alpha_1}\alpha_2^{-1}
\end{eqnarray}
